---
output: github_document
number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=7.5)

library(data.table)
library(sp)
library(ggplot2)
library(scico)

library(ggpubr)

gfcov <- readRDS(paste0("~/floodGAM/data/processed-data/","gamfelt/",
                       "gamfelt_catchment_covariates.rds"))
load(paste0("~/floodGAM/data/processed-data/gamfelt/README_files/stationmaps.rda"))


percentfin <- readRDS(paste0("~/floodGAM/data/","README_files/",
                                "findata_stats.rds"))
```

# The `gamfelt` dataset

This folder contains the annual maxima and catchment descriptors for the `gamfelt` dataset.

## File structure

 - `gamfelt_1hour_annual_maxima` - the annual maxima, in $m^3/s$. Data from archive hyfin-complete. Interpolated to hourly values and annual maxima extracted (see processing scripts [here](~/floodGAM/code/scripts/1-data-creation/). Saved as both `.csv` and `.rds`.
 - `gamfelt_catchment_covariates` - catchment descriptors for each gamfelt station. Saved as both `.csv` and `.rds`.
 - `gamfelt_catchment_covariates_description.txt` - description of each variable in the catchment descriptor dataset.


## Describing the dataset

The `gamfelt` dataset contains annual maxima for 241 stations in Norway. Each station has at least 20 years of fine data.

The stations represent a reasonably diverse collection of catchment sizes and hydroclimatic regimes:

```{r,echo=F,warning=F,message=F,fig.height=5}
figure <- ggarrange(gprecip, gstations, gtemperature,
                    widths=c(1,1,1),
                    nrow=1)

figure
```


Total record lengths range from a minimum of 20 years of data to around 60 years of fine (sub-daily) data.

```{r,echo=F,fig.dim=c(5,4)}
gfam <- readRDS(paste0("~/floodGAM/data/processed-data/gamfelt-durations/",
                       "gamfelt_durations_annual_maxima.rds"))

hist(gfam[d==1,.N,by="ID"]$N,breaks=15,
     xlab="Record length (years)",ylab="Stations",main="",
     col="white")
```

